Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-06-18 sports_analytics.Rmd
wp in football now looks decent

In recent years, a number of statistical methods have been used to build well-calibrated win probability models in American football. @Lock2014 used random forests as a means for estimating win probability before each play in a football game. In addition, @Yurko2019 estimated win probability in the NFL using a generalized additive model, as part of the `nflscrapr` package and nflwar framework. A major highlight of this framework is it is fully reproducible and uses publicly available data. One of @Yurko2019's goals was also to encourage researchers to "use, explore, and improve upon our work", which ultimately inspired \pkg{nflfastR}, known as the successor to `nflscrapr`

(add a wp example from nflfastR)

@Yam2019 causal

@Yurko2020 going deep

@goldner2017situational

>>>>

In recent years, a number of statistical methods have been used to build well-calibrated win probability models in American football. These are flexible models that have high predictability, can account for nonlinear interactions between the explanatory variables, require little assumptions, and produce feature importance scores.

@Lock2014 used random forests as a means for estimating win probability before each play in a football game. Covariates included in this win probability model are the current down, score differential, time remaining, adjusted score, point spread, number of timeouts remaining for each team, total points scored, current yardline, and yards to go for a first down.

In addition, @Yurko2019 estimated win probability in the NFL using a generalized additive model (GAM), as part of the `nflscrapr` package and nflWAR framework. This model takes into account the estimated expected points obtained from the model described in \ref{sec:ev}, along with other predictors for time, current half, and timeouts. A major highlight of this framework is it is fully reproducible and uses publicly available data. One of @Yurko2019's goals was also to encourage researchers to "use, explore, and improve upon our work", which ultimately inspired \pkg{nflfastR}, known as the successor to `nflscrapr`

These two frameworks were also implemented in @Yam2019 with minimal modifications. Specifically, the authors combined both @Lock2014 and @Yurko2019's approaches to estimate the win probability for each play, with an overall goal of assessing fourth down decision making in gridiron football.

(add a wp example from nflfastR or nflscapr)
